Mathematical Language Processing \\ Project

### 2014-06-17 paper.tex
Sec 2 is not unerstandable

In natural language, words and phrases themselves imply the semantics. In
contrast, the meaning of identifiers in mathematical formulae is undefined.
Thus scientists must study the context to decode the meaning. The
\emph{Mathematical Language Processing} project aims to support that process.
In this paper, we compare two approaches to discover identifier - definition tuples.
At first we use a simple pattern matching
approach. Second, we present the \emph{MLP} approach that uses POS tag based
distances as well as sentence positions to calculate identifier - definition
probabilities.
The evaluation of our prototypical system, applied on the
Wikipedia text corpus, shows that our approach augments the user experience
substantially. While hovering the identifiers in the formula, tool-tips with
the most probable definitions occur. Tests with random samples show that the
displayed definitions provide a good match with the actual meaning of the
identifiers.

@@

To build such a system, a labeled text corpus that annotates identifiers and
their definition is desirable. At the project start, such a corpus was not
available. We use simple statistical assumptions about co-occurrences between
identifiers and their potential description to extract the meaning of the
identifiers from their context. In combination with the Stratosphere system
that is capable of scalable data processing,  we rapidly generate a large
amount of definition relation candidates with only minimal implementation
overhead to ensure the scalability of our algorithm.

We chose the Wikipedia text corpus as a target because of two facts. First,
most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
language) for formulae. It is easy to detect and the extraction of identifiers
is well tried. Second, the text is already annotated with mark-up.
Particularly, hyperlinks to other articles within the Wikipedia are of
interest as they typically wrap around any number of words and indicate that
these in combination are relevant in the given context (respectively
sentence).

@@

At first, we implemented a simple identifier definition extractor that is
based on a set of static patterns. As it is an fairly robust approach und easy
to implement, this will serve us as a good reference point in terms of
performance. The following patterns are being used to discover description
terms:

@@

Note that in the last pattern we used a \emph{part-of-speech} tag. Actually
determiners not only contain articles but also quatifiers and distributives.
To increase readability this list have been shortened to `*/DT'.

>>>>

In natural language, words and phrases themselves imply the semantics. In contrast, the meaning of identifiers in mathematical formulae is undefined.
Thus scientists must study the context to decode the meaning. The \emph{Mathematical Language Processing} (MLP) project aims to support that process.  In this paper, we compare two approaches to discover identifier-definition tuples.  At first we use a simple pattern matching approach. Second, we present the \emph{MLP} approach that uses Part-of-speech (POS) tag based distances as well as sentence positions to calculate indentifier-definition probabilities. The evaluation of our prototypical system, applied on the Wikipedia text corpus, shows that our approach augments the user experience substantially. While hovering the identifiers in the formula, tool-tips with the most probable definitions occur. Tests with random samples show that the displayed definitions provide a good match with the actual meaning of the identifiers.

@@

To build such a system, a labelled text corpus that annotates identifiers and
their definition is desirable. At the project start, such a corpus was not
available. 
We assume that the probability to find correct identifier definition tuples is
POS tagged based distance metrics.
Our motivation for that is readers usually extract the identifier definition
from the context that is given by the surrounding text.
In combination with the Stratosphere system
that is capable of scalable data processing,  we rapidly generate a large
amount of definition relation candidates with only minimal implementation
overhead for the parallelization.

We chose the Wikipedia text corpus as a target because of two facts. First,
most articles make use of \texttt{<math/>} tags (\emph{texvc} as an input
language) for formulae. Those tags are easy to detect and the extraction identifiers
in its experimental MathML output format is well tried. 
Second, the text is already annotated with mark-up.
Particularly, hyperlinks to other articles within the Wikipedia are of
interest as they typically wrap around any number of words and indicate that
these in combination are relevant in the given context (respectively
sentence).

@@

At first, we implemented a simple identifier definition extractor that is
based on a set of static patterns. As it is an fairly robust approach and easy
to implement, this will serve us as a good reference point in terms of
performance. The following patterns are being used to discover description
terms:

@@

Note that in the last pattern we used a \emph{part-of-speech} tag. Actually
determiners not only contain articles but also quantifiers and distributives.
To increase readability this list have been shortened to `*/DT'.
\todo[inline]{describe the process: extraction of identifiers, how to didentify the dscription? Looking for space or using pos tags?}
